---
title: "Oridinal Logistic Regression"
author: "Ian Dinwoodie"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(MASS)
library(AER)
library(plyr)
#library(nnet)
#library(rms)
#library(VGAM)
library(brant)
library(ordinal)
opts_chunk$set(echo=TRUE)
```

# Preparing the Data

Load the dataframe and check the dimensions. We expect 963 rows and 301 columns.

```{r}
# Load the data from the csv into a data frame.
df.orig <- read.csv("data/processed.csv", header=TRUE)

# Get the original data frame dimensions.
dim(df.orig)
```

Some columns need to have their values standardized.

```{r}
# Convert "dog_sex" into "male" columns.
df.mod <- df.orig
df.mod$male <- ifelse(df.mod$dog_sex==1, 1, 0)
df.mod <- df.mod[!is.na(df.mod$male),]

# Create dummy variables for the levels 1-4 of "behavior_prog".
# We ignore levels 5 and 6 because they are irrelevant to the question.
df.mod$behavior_prog_1 <- ifelse(df.mod$behavior_prog==1, 1, 0)
df.mod$behavior_prog_1[is.na(df.mod$behavior_prog_1)] <- 0
df.mod$behavior_prog_1 <- as.integer(df.mod$behavior_prog_1)
df.mod$behavior_prog_2 <- ifelse(df.mod$behavior_prog==2, 1, 0)
df.mod$behavior_prog_2[is.na(df.mod$behavior_prog_2)] <- 0
df.mod$behavior_prog_2 <- as.integer(df.mod$behavior_prog_2)
df.mod$behavior_prog_3 <- ifelse(df.mod$behavior_prog==3, 1, 0)
df.mod$behavior_prog_3[is.na(df.mod$behavior_prog_3)] <- 0
df.mod$behavior_prog_3 <- as.integer(df.mod$behavior_prog_3)
df.mod$behavior_prog_4 <- ifelse(df.mod$behavior_prog==4, 1, 0)
df.mod$behavior_prog_4[is.na(df.mod$behavior_prog_4)] <- 0
df.mod$behavior_prog_4 <- as.integer(df.mod$behavior_prog_4)

# Add a column indicating if the dog was found to have pain.
df.mod$pain <- ifelse(df.mod$health_iss_3==1, 1, 0)

# Add a column indicating if the dog was found to have fear/anxiety.
df.mod$fear_anxiety <- ifelse(df.mod$q02_main_2==1, 1, 0)

# Correct training method entries.
# We consider dogs trained with any form of punishment to be excluded from a
# reward-only training style. We consider compulsion a form of punishment.
df.mod$punish <- ifelse(
    ((df.mod$training_methods_2==1) | (df.mod$training_methods_3==1)
     | (df.mod$training_methods_3==1)), 1, 0)
df.mod$reward <- ifelse(df.mod$punish==1, 0, df.mod$training_methods_1)
df.mod <- subset(df.mod,
             select=-c(training_methods_1, training_methods_2,
                       training_methods_3,training_methods_4))

# Drop clicker/whistle from equipment since it is present in the behavior
# techniques.
df.mod <- subset(df.mod, select=-equipment_11)

# Define bite severity levels as follows:
# L1 = Has not bitten.
# L2 = Has bitten, but not punctured skin.
# L3 = Has bitten and punctured skin.
#names(df.mod)[names(df.mod)=='q03_severity'] <- 'bite_severity'
#df.mod$bite_severity <- ifelse(
#    is.na(df.mod$bite_severity), df.mod$q03_form_5, df.mod$bite_severity)
#df.mod$bite_severity <- ifelse(df.mod$bite_severity>2, 3, df.mod$bite_severity)
#df.mod$bite_severity <- ifelse(df.mod$bite_severity<1, 1, df.mod$bite_severity)
#df.mod$bite_severity <- factor(
#    df.mod$bite_severity, levels=c("1","2","3"), ordered=TRUE)

names(df.mod)[names(df.mod)=='q03_form_5'] <- 'has_bitten'
df.mod$has_bitten <- as.factor(df.mod$has_bitten)

df.mod$trainer_cred <- ifelse(df.mod$trainer_cred == 1, 1, 0)
df.mod[is.na(df.mod$trainer_cred), 'trainer_cred'] <- 0
names(df.mod)[names(df.mod)=='trainer_cred_type'] <- 'trainer_cpdt'
df.mod$trainer_cpdt <- ifelse(df.mod$trainer_cpdt == 1, 1, 0)
df.mod[is.na(df.mod$trainer_cpdt), 'trainer_cpdt'] <- 0

df.mod$consultant_cred <- ifelse(df.mod$consultant_cred == 1, 1, 0)
df.mod[is.na(df.mod$consultant_cred), 'consultant_cred'] <- 0
names(df.mod)[names(df.mod)=='consultant_cred_type_1'] <- 'consult_cbcc'
df.mod$consult_cbcc <- ifelse(df.mod$consult_cbcc == 1, 1, 0)
names(df.mod)[names(df.mod)=='consultant_cred_type_2'] <- 'consult_cdbc'
df.mod$consult_cdbc <- ifelse(df.mod$consult_cdbc == 1, 1, 0)
names(df.mod)[names(df.mod)=='consultant_cred_type_3'] <- 'consult_caab'
df.mod$consult_caab <- ifelse(df.mod$consult_caab == 1, 1, 0)
names(df.mod)[names(df.mod)=='consultant_cred_type_4'] <- 'consult_dacvb'
df.mod$consult_dacvb <- ifelse(df.mod$consult_dacvb == 1, 1, 0)

str(df.mod)
```

Retain only the columns necessary for analysis.

```{r}
# Retain only the columns to be used for analysis.
predictors <- c(
  '^male$',
  '^prof_type_\\d$',
  '^equipment_\\d+$',
  '^med_list_\\d+$',
  '^alt_med_type_\\d+$',
  '^behavior_prog_\\d+$',
  '^behavior_tech_used_\\d+$',
  '^pain$',
  '^fear_anxiety$',
  '^reward$',
  '^has_bitten$',
  'trainer_cred',
  'trainer_cpdt',
  '^consultant_cred$',
  '^consult_cbcc$',
  '^consult_cdbc$',
  '^consult_caab$',
  '^consult_dacvb$'
)
outcomes <- c(
  "^agg_\\S+$"
)
pattern <- paste(c(predictors, outcomes), collapse='|')
idx <- grep(pattern, names(df.mod))
df.mod <- df.mod[, idx]
dim(df.mod)
```

All remaining columns except for the response to treatment for aggression scores
are factors. Update the column types to reflect this.

```{r}
for (col in names(df.mod)) {
  if (grepl("agg_", col, fixed=TRUE)) next
  df.mod[, col] <- as.factor(df.mod[, col])
}

str(df.mod)
```

General rule of thumb for logistic regression is that each variable should have
at least 10 responses per response option. We apply this rule of thumb below.

```{r}
# Drop the columns that do not meet the minimum response cutoff.
apply_binary_response_criteria <- function(df)
{
  min_response_cutoff <- 10
  drops <- NULL
  for (col in names(df)) {
    if (grepl("agg_", col, fixed=TRUE)) next
    counts <- count(df[, col])
    if (nrow(counts) < 2) {
      drops <- c(drops, col)
      break
    }
    for (row in 1:nrow(counts)) {
      if (counts[row, "freq"] < min_response_cutoff) {
        drops <- c(drops, col)
        break
      }
    }
  }
  
  return(df[, !(names(df) %in% drops)])
}

df.mod <- apply_binary_response_criteria(df.mod)
summary(df.mod)
```

Now we initialize other functions that will be reused throughout the analysis.

```{r}
get_agg_data_frame <- function(df, pred.patterns, outcome, excludes)
{
  pattern <- paste(c(pred.patterns, outcome), collapse='|')
  idx <- grep(pattern, names(df))
  df.out <- df[, idx]
  
  # Format the outcome as a factor with order.
  df.out[,outcome] <- round(df.out[, outcome], 0)
  df.out[,outcome] <- factor(
      df.out[, outcome], levels=c("1","2","3","4","5","6","7"), ordered=TRUE)
  
  if (length(excludes) > 0) {
    # Exclude variables that cause issues for the model.
    df.out <- df.out[, !colnames(df.out) %in% excludes]
  }
  
  # Drop rows with null values for the outcome.
  df.out <- df.out[!is.na(df.out[, outcome]), ]
  
  return(df.out)
}
```

# Overall Aggression

```{r polr_model}
set.seed(1)
iters <- 10

# Generate the desired subset data frame.
outcome <- "agg_avg"
excludes <- vector()
df.agg <- apply_binary_response_criteria(
    get_agg_data_frame(df.mod, predictors, outcome, excludes))
summary(df.agg)

# Fit a logistic regression.
f <- as.formula(paste0(outcome, "~", "."))
coef <- rep(0, length(df.agg)-1)
zeta <- seq(from=0, to=1, length.out=6)
start <- c(coef, zeta)
m <- polr(f, data=df.agg, Hess=TRUE, start=start)
summary(m)

fit2 <- lm(as.numeric(agg_avg) ~ ., data=df.agg)
print(vif(fit2))

# Generate CIs and p-values.
coeftest(m) # p-values
ci <- exp(confint.default(m))
odds.ratio <- exp(coef(m))
cbind(odds.ratio, ci) # OR and CI
```

```{r clm_model}
fm <- ordinal::clm(agg_avg ~ has_bitten + prof_type_3 + prof_type_1 + prof_type_2 + trainer_cred + trainer_cpdt + consultant_cred + consult_cbcc + consult_cdbc + consult_caab + consult_dacvb + equipment_1 + equipment_2 + equipment_3 + equipment_4 + equipment_5 + equipment_9 + equipment_6 + equipment_7 + equipment_10 + equipment_8 + med_list_1 + med_list_7 + med_list_9 + med_list_10 + med_list_11 + med_list_13 + alt_med_type_1 + alt_med_type_2 + alt_med_type_3 + alt_med_type_4 + alt_med_type_6 + alt_med_type_8 + alt_med_type_11 + alt_med_type_12 + alt_med_type_13 + alt_med_type_15 + alt_med_type_16 + alt_med_type_18 + behavior_tech_used_1 + behavior_tech_used_3 + behavior_tech_used_4 + behavior_tech_used_5 + behavior_tech_used_6 + behavior_tech_used_7 + behavior_tech_used_8 + behavior_tech_used_9 + behavior_tech_used_10 + behavior_tech_used_11 + behavior_tech_used_12 + behavior_tech_used_13 + behavior_tech_used_15 + behavior_tech_used_16 + behavior_tech_used_17 + behavior_tech_used_18 + behavior_tech_used_19 + male + behavior_prog_1 + behavior_prog_2 + behavior_prog_3 + behavior_prog_4 + pain + fear_anxiety + reward + behavior_tech_used_2, data=df.agg)
summary(fm)
```
```{r nominal_test, warning=FALSE}
nominal_test(fm)
#scale_test(fm)
```

We see that the test fails for three of the variables. Now let's check what
happens if we specify those variables as nominal effects.

```{r clm_model_2}
fm2 <- ordinal::clm(agg_avg ~ has_bitten + prof_type_3 + prof_type_1 + prof_type_2 + trainer_cred + trainer_cpdt + consultant_cred + consult_cbcc + consult_cdbc + consult_caab + consult_dacvb + equipment_1 + equipment_2 + equipment_3 + equipment_4 + equipment_5 + equipment_9 + equipment_6 + equipment_7 + equipment_10 + equipment_8 + med_list_1 + med_list_7 + med_list_9 + med_list_10 + med_list_11 + med_list_13 + alt_med_type_1 + alt_med_type_2 + alt_med_type_3 + alt_med_type_4 + alt_med_type_6 + alt_med_type_8 + alt_med_type_11 + alt_med_type_12 + alt_med_type_13 + alt_med_type_15 + alt_med_type_16 + alt_med_type_18 + behavior_tech_used_1 + behavior_tech_used_3 + behavior_tech_used_4 + behavior_tech_used_5 + behavior_tech_used_6 + behavior_tech_used_7 + behavior_tech_used_8 + behavior_tech_used_9 + behavior_tech_used_10 + behavior_tech_used_11 + behavior_tech_used_12 + behavior_tech_used_13 + behavior_tech_used_15 + behavior_tech_used_16 + behavior_tech_used_17 + behavior_tech_used_18 + behavior_tech_used_19 + male + behavior_prog_1 + behavior_prog_2 + behavior_prog_3 + behavior_prog_4 + pain + fear_anxiety + reward,
                    nominal = ~ behavior_tech_used_2 + behavior_tech_used_1
                        + behavior_tech_used_4,
                    data=df.agg,
                    threshold = "equidistant")
summary(fm2)
```
```{r nominal_test_2, warning=FALSE}
nominal_test(fm2)
#scale_test(fm2)
```

Now even more variables are failing the parallel line assumption. Let's go back
to our original model and collapse the 

Let's look at a binomial regression instead.

```{r binomial_model}
df_agg2 <- df.agg
df_agg2$agg_avg <- ifelse(df_agg2$agg_avg <= 4, 0, 1)

fm2 <- glm(agg_avg ~ has_bitten + prof_type_3 + prof_type_1 + prof_type_2 + trainer_cred + trainer_cpdt + consultant_cred + consult_cbcc + consult_cdbc + consult_caab + consult_dacvb + equipment_1 + equipment_2 + equipment_3 + equipment_4 + equipment_5 + equipment_9 + equipment_6 + equipment_7 + equipment_10 + equipment_8 + med_list_1 + med_list_7 + med_list_9 + med_list_10 + med_list_11 + med_list_13 + alt_med_type_1 + alt_med_type_2 + alt_med_type_3 + alt_med_type_4 + alt_med_type_6 + alt_med_type_8 + alt_med_type_11 + alt_med_type_12 + alt_med_type_13 + alt_med_type_15 + alt_med_type_16 + alt_med_type_18 + behavior_tech_used_1 + behavior_tech_used_3 + behavior_tech_used_4 + behavior_tech_used_5 + behavior_tech_used_6 + behavior_tech_used_7 + behavior_tech_used_8 + behavior_tech_used_9 + behavior_tech_used_10 + behavior_tech_used_11 + behavior_tech_used_12 + behavior_tech_used_13 + behavior_tech_used_15 + behavior_tech_used_16 + behavior_tech_used_17 + behavior_tech_used_18 + behavior_tech_used_19 + male + behavior_prog_1 + behavior_prog_2 + behavior_prog_3 + behavior_prog_4 + pain + fear_anxiety + reward + behavior_tech_used_2 + behavior_tech_used_1 + behavior_tech_used_4,
                    data=df_agg2, family="binomial")
summary(fm2)
```

```{r binomial_confint}
confint(fm2)
```

# Conflict Aggression

```{r}
set.seed(1)

# Generate the desired subset data frame.
outcome <- "agg_conf"
df.agg <- apply_binary_response_criteria(
    get_agg_data_frame(df.mod, predictors, outcome, excludes))
summary(df.agg)

# Fit a logistic regression.
f <- as.formula(paste0(outcome, "~", "."))
coef <- rep(0, length(df.agg)-1)
zeta <- seq(from=0, to=1, length.out=6)
start <- c(coef, zeta)
m <- polr(f, data=df.agg, Hess=TRUE, start=start)
summary(m)

fit2 <- lm(as.numeric(agg_conf) ~ ., data=df.agg)
print(vif(fit2))

# Generate CIs and p-values.
coeftest(m) # p-values
ci <- exp(confint.default(m))
odds.ratio <- exp(coef(m))
cbind(odds.ratio, ci) # OR and CI
```

# I.D.H. Aggression

```{r}
# Generate the desired subset data frame.
outcome <- "agg_idh"
df.agg <- apply_binary_response_criteria(
  get_agg_data_frame(df.mod, predictors, outcome, excludes))
summary(df.agg)

# Fit a logistic regression.
f <- as.formula(paste0(outcome, "~", "."))
coef <- rep(0, length(df.agg)-1)
zeta <- seq(from=0, to=1, length.out=6)
start <- c(coef, zeta)
m <- polr(f, data=df.agg, Hess=TRUE, start=start)
summary(m)

fit2 <- lm(as.numeric(agg_idh) ~ ., data=df.agg)
print(vif(fit2))

# Generate CIs and p-values.
coeftest(m) # p-values
ci <- exp(confint.default(m))
odds.ratio <- exp(coef(m))
cbind(odds.ratio, ci) # OR and CI
```

# External Aggression (People)

```{r}
# Generate the desired subset data frame.
outcome <- "agg_ext_ppl"
df.agg <- apply_binary_response_criteria(
    get_agg_data_frame(df.mod, predictors, outcome, excludes))
summary(df.agg)

# Fit a logistic regression.
f <- as.formula(paste0(outcome, "~", "."))
coef <- rep(0, length(df.agg)-1)
zeta <- seq(from=0, to=1, length.out=6)
start <- c(coef, zeta)
m <- polr(f, data=df.agg, Hess=TRUE, start=start)
summary(m)

fit2 <- lm(as.numeric(agg_ext_ppl) ~ ., data=df.agg)
print(vif(fit2))

# Generate CIs and p-values.
coeftest(m) # p-values
ci <- exp(confint.default(m))
odds.ratio <- exp(coef(m))
cbind(odds.ratio, ci) # OR and CI
```

# External Aggression (Dogs)

```{r}
# Generate the desired subset data frame.
outcome <- "agg_ext_dog"
df.agg <- apply_binary_response_criteria(
    get_agg_data_frame(df.mod, predictors, outcome, excludes))
summary(df.agg)

# Fit a logistic regression.
f <- as.formula(paste0(outcome, "~", "."))
print(f)
coef <- rep(0, length(df.agg)-1)
zeta <- seq(from=0, to=1, length.out=6)
start <- c(coef, zeta)
m <- polr(f, data=df.agg, Hess=TRUE, start=start)
summary(m)

fit2 <- lm(as.numeric(agg_ext_dog) ~ ., data=df.agg)
print(vif(fit2))

# Generate CIs and p-values.
coeftest(m) # p-values
ci <- exp(confint.default(m))
odds.ratio <- exp(coef(m))
cbind(odds.ratio, ci) # OR and CI
```

# Predatory Aggression

```{r}
# Generate the desired subset data frame.
outcome <- "agg_pred"
df.agg <- apply_binary_response_criteria(
    get_agg_data_frame(df.mod, predictors, outcome, excludes))
summary(df.agg)

# Fit a logistic regression.
f <- as.formula(paste0(outcome, "~", "."))
coef <- rep(0, length(df.agg)-1)
zeta <- seq(from=0, to=1, length.out=6)
start <- c(coef, zeta)
m <- polr(f, data=df.agg, Hess=TRUE, start=start)
summary(m)

fit2 <- lm(as.numeric(agg_pred) ~ ., data=df.agg)
print(vif(fit2))

# Generate CIs and p-values.
coeftest(m) # p-values
ci <- exp(confint.default(m))
odds.ratio <- exp(coef(m))
cbind(odds.ratio, ci) # OR and CI
```
