---
title: "Types of Professionals"
author: "Ian Dinwoodie"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(epitools)
opts_chunk$set(echo=TRUE)
set.seed(1)
```

# Analysis for Types of Professionals

## Preparing the Data Frame

Load the data frame and check the dimensions. We expect the data frame to have
963 rows and 298 columns. We then drop all the columns that are not necessary
for calculations pertaining to the types of professionals sought.

```{r}
# Load the data from the csv into a data frame.
df <- read.csv("data/processed.csv", header=TRUE)
print(dim(df))

# Retaining only the necessary columns.
attribs <- c(
  'prof_help',
  'prof_type_*',
  'prof_discover',
  'health_iss_*',
  'trainer_*',
  'consultant_*',
  'best_prof*',
  'alt_med_type_4'
)
outcomes <- c('agg_\\.*')
keep.patterns <- paste(c(attribs, outcomes), collapse='|')
df <- df[, grep(keep.patterns, names(df))]

# Display the remaining columns.
str(df)
```

With a more managable sized data frame we can rename the columns.

```{r}
# Rename the columns for accessibility.
names <- c(
  'prof.help',
  'vet',
  'trainer',
  'consult',
  'health.issue',
  'thyroid.problem',
  'seizures',
  'pain',
  'tumor',
  'cognitive.dysfunction',
  'other.health.issue',
  'trainer.multi',
  'trainer.cnt',
  'trainer.has.cred',
  'trainer.cpdt',
  'trainer.cred.other.desc',
  'consult.multi',
  'consult.cnt',
  'consult.has.cred',
  'consult.cbcc',
  'consult.cdbc',
  'consult.caab',
  'consult.dacvb',
  'consult.other',
  'consult.cred.other.desc',
  'prof.best',
  'chiro',
  'agg.conf',
  'agg.idh',
  'agg.ext.ppl',
  'agg.ext.dog',
  'agg.pred',
  'agg.overall'
)
colnames(df) <- names

# Peek at the structure of the data.
str(df)
```

Now let's convert the columns into appropriate types.

```{r}
# Convert columns to appropriate type.
df$prof.help <- ifelse(df$prof.help == 1, TRUE, FALSE)
df$vet <- ifelse(df$vet == 1, TRUE, FALSE)
df$trainer <- ifelse(df$trainer == 1, TRUE, FALSE)
df$consult <- ifelse(df$consult == 1, TRUE, FALSE)
df$health.issue <- ifelse(df$health.issue == 1, TRUE, FALSE)
df$thyroid.problem <- ifelse(df$thyroid.problem == 1, TRUE, FALSE)
df$seizures <- ifelse(df$seizures == 1, TRUE, FALSE)
df$pain <- ifelse(df$pain == 1, TRUE, FALSE)
df$tumor <- ifelse(df$tumor == 1, TRUE, FALSE)
df$cognitive.dysfunction <- ifelse(df$cognitive.dysfunction == 1, TRUE, FALSE)
df$other.health.issue <- ifelse(df$other.health.issue == 1, TRUE, FALSE)
df$trainer.multi <- ifelse(df$trainer.multi == 1, TRUE, FALSE)
#df$trainer.cnt <- as.factor(df$trainer.cnt)
df$trainer.has.cred <- ifelse(df$trainer.has.cred == 1, TRUE, FALSE)
df$trainer.cpdt <- ifelse(df$trainer.cpdt == 1, TRUE, FALSE)
#df$trainer.cred.other.desc <- ifelse(df$trainer.cred.other)
df$consult.multi <- ifelse(df$consult.multi == 1, TRUE, FALSE)
#df$consult.cnt <- as.factor(df$consult.cnt)
df$consult.has.cred <- ifelse(df$consult.has.cred == 1, TRUE, FALSE)
df$consult.cbcc <- ifelse(df$consult.cbcc == 1, TRUE, FALSE)
df$consult.cdbc <- ifelse(df$consult.cdbc == 1, TRUE, FALSE)
df$consult.caab <- ifelse(df$consult.caab == 1, TRUE, FALSE)
df$consult.dacvb <- ifelse(df$consult.dacvb == 1, TRUE, FALSE)
df$consult.other <- ifelse(df$consult.other == 1, TRUE, FALSE)
#df$consult.cred.other.desc <- ifelse
df$prof.best <- as.factor(df$prof.best)
#df$agg.conf
#df$agg.idh
#df$agg.ext.ppl
#df$agg.ext.dog
#df$agg.pred
#df$agg.overall
```

## Overview

Let's take a look at a summary of the data frame.

```{r}
# Summarize the data frame.
summary(df)
```

Let's look specifically at the professionals that participants marked as best
(i.e., most helpful) since it isn't fully expanded in the summary above.

```{r}
summary(df$best.prof)
```

Clearly define the outcome variables for this analysis.

```{r}
outcomes <- c('agg.overall', 'agg.conf', 'agg.idh', 'agg.ext.ppl',
              'agg.ext.dog', 'agg.pred')
```

We also establish some functions that will be used throughout the course of this
analysis.

```{r}
iters <- 20

get_outcome_df <- function(df, outcomes, attribs, response.min=5)
{
  # Reduce the data frame.
  cols <- c(outcome, attribs)
  df.out <- df[cols]
  
  # Format the outcome as a factor with order.
  df.out[,outcome] <- round(df.out[,outcome], 0)
  df.out[,outcome] <- factor(
    df.out[,outcome], levels=c("1","2","3","4","5","6","7"), ordered=TRUE)

  # Drop rows with null values for the outcome.
  df.out <- df.out[!is.na(df.out[,outcome]), ]
  
  # Each attribute must meet the minimum response cutoff.
  drop.idxs <- NULL
  attr.idxs <- grep(attribs, names(df.out))
  response.max <- nrow(df.out)
  for (idx in attr.idxs) {
    for (c in table(df.out[, idx])) {
      if (c<=(response.min-1) || c>=response.max) {
        drop.idxs <- c(drop.idxs, idx)
        break
      }
    }
  }
  
  # Drop the columns that did not meet the minimum response cutoff.
  if (length(drop.idxs) > 0) {
    df.out <- df.out[, -drop.idxs]
  }
  
  return(df.out)
}

get_chi_square_results <- function(df, outcome, attribs)
{
  cat(paste("\n", outcome, "\n"))
  df1 <- get_outcome_df(df, outcome, attribs, response.min=5)
  
  # Format the outcome as a binary response.
  outcome.binary <- paste0(outcome, ".improve")
  df1[, outcome.binary] <- ifelse(df1[, outcome] > 4, 1, 0)
  
  # Get p-values and odds ratios.
  preds <- NULL
  p.values <- NULL
  odds.ratios <- NULL 
  ci.low <- NULL
  ci.high <- NULL
  for (col in names(df1)) {
    if (col == outcome || col == outcome.binary) next
    set.seed(1)
    
    tbl <- table(df1[, col], df1[, outcome.binary], dnn=c(col, outcome.binary))
    chisq <- chisq.test(tbl, simulate.p.value=TRUE)
    p <- chisq$p.value
    
    preds <- c(preds, col)
    p.values <- c(p.values, p)
    or <- oddsratio.wald(df1[, col], df1[, outcome.binary])
    odds.ratios <- c(odds.ratios, or$measure[2,1])
  
    b.ors <- NULL
    for (i in 1:iters) {
      df.tmp <- df1[sample(nrow(df1), nrow(df), replace=T), ]
      or <- oddsratio.wald(df.tmp[, col], df.tmp[, outcome.binary])
      b.ors <- c(b.ors, or$measure[2,1])
    }
    b.ors <- b.ors[is.finite(b.ors)]
    b.ors <- b.ors[b.ors != 0]
    
    ci <- quantile(b.ors, c(0.025, 0.975), na.rm=TRUE)
    ci.low <- c(ci.low, ci[1])
    ci.high <- c(ci.high, ci[2])
  }

  # Form a result data frame.
  df.out <- data.frame(pred=preds,
                       p=round(p.values, 5),
                       or=round(odds.ratios, 4),
                       ci.low=as.vector(round(ci.low, 4)),
                       ci.high=as.vector(round(ci.high, 4)))
  
  df.out$sig <- ''
  #df.out[df.out$p <= .1, 'sig'] <- '.'
  df.out[df.out$p <= .05, 'sig'] <- '*'
  df.out[df.out$p <= .01, 'sig'] <- '**'
  df.out[df.out$p <= .001, 'sig'] <- '***'
  
  for (i in 1:nrow(df.out)) {
    if (is.na(df.out[i, 'ci.low']) | is.na(df.out[i, 'ci.high'])) next
    if ((df.out[i, 'ci.low'] < 1) & (df.out[i, 'ci.high'] > 1)) {
      df.out[i, 'sig'] <- 'x'
    }
  }

  df.out$dir <- ''
  df.out[df.out$or < 1, 'dir'] <- '-'
  df.out[df.out$or > 1, 'dir'] <- '+'
  
  return(df.out)
}
```

## Trainers

First, we look at the whole sample to see if seeking out a trainer had any
correlation with a positive response to treatment for aggression (PRTA).

```{r}
# Specify the predictor columns.
attribs <- c('trainer')

# Run chi square analysis.
for (outcome in outcomes) {
  df.out = get_chi_square_results(df, outcome, attribs)
  print(df.out)
}
```

Now we separate out the group of dogs that were brought to trainers and look for
any correlations between a trainer having credentials and a PRTA.

```{r}
# Isolate the group of dogs that saw trainers. We expect to have 384 rows in the
# resulting data frame.
df.tmp <- subset(df, trainer == TRUE)
dim(df.tmp)

# Specify the predictor columns.
attribs <- c('trainer.has.cred')

# Run chi square analysis.
for (outcome in outcomes) {
  df.out = get_chi_square_results(df, outcome, attribs)
  print(df.out)
}
```

Next, we separate out the trainers with credentials and look for any
correlations between specific trainer credentials and a PRTA.

```{r}
# Isolate the group of dogs that saw trainers. We expect to have 274 rows in the
# resulting data frame.
df.tmp <- subset(df.tmp, trainer.has.cred == TRUE)
dim(df.tmp)

# Specify the predictor columns.
attribs <- c('trainer.cpdt')

# Run chi square analysis.
for (outcome in outcomes) {
  df.out = get_chi_square_results(df, outcome, attribs)
  print(df.out)
}
```

## Behavior Consultants

We look at the whole sample to see if seeking out a behavior consultant had any
correlation with a PRTA.

```{r}
# Specify the predictor columns.
attribs <- c('consult')

# Run chi square analysis.
for (outcome in outcomes) {
  df.out = get_chi_square_results(df, outcome, attribs)
  print(df.out)
}
```

Now we separate out the group of dogs that were brought to consultants and look
for any correlations between a consultant having credentials and a PRTA.

```{r, warning=F}
# Isolate the group of dogs that saw consultants. We expect to have 276 rows in
# the resulting data frame.
df.tmp <- subset(df, consult == TRUE)
dim(df.tmp)

# Specify the predictor columns.
attribs <- c('consult.has.cred')

# Run chi square analysis.
for (outcome in outcomes) {
  df.out = get_chi_square_results(df, outcome, attribs)
  print(df.out)
}
```

Next, we separate out the consultants with credentials and look for any
correlations between specific consultant credentials and a PRTA.

```{r, warning=F}
# Isolate the group of dogs that saw consultants. We expect to have 196 rows in
# the resulting data frame.
df.tmp <- subset(df.tmp, consult.has.cred == TRUE)
dim(df.tmp)

# Specify the predictor columns.
attribs <- c('consult.cbcc', 'consult.cdbc', 'consult.caab', 'consult.dacvb')

# Run chi square analysis.
for (outcome in outcomes) {
  df.out = get_chi_square_results(df, outcome, attribs)
  print(df.out)
}
```

## Veterinarians

We look at the whole sample to see if seeking out a veterinarian had any
correlation with a PRTA.

```{r}
# Specify the predictor columns.
attribs <- c('vet')

# Run chi square analysis.
for (outcome in outcomes) {
  df.out = get_chi_square_results(df, outcome, attribs)
  print(df.out)
}
```

We would also like to see how many dogs with pain exhibit fear aggression toward
dogs.

```{r}
# Pain & FAD
df.tmp <- subset(df, (pain == TRUE) & (agg.ext.dog > 0))
print(summary(df.tmp))

# Pain & chiro
df.tmp <- subset(df, (pain == TRUE) & (chiro > 0))
print(summary(df.tmp))
```